---
title: "Final Project"
author: "Rachael Cooper, Sarah Gould, Nathan Patton, Samarth Saxena"
date: "11/29/2021"
output: 
  html_document:
    toc: TRUE
    toc_depth: 5
    toc_float: TRUE
    tod_collapsed: TRUE
    theme: cosmo
  editor_options:
    chunk_output_type: console
    
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# cache --> only have to recompile the changes
# eval --> runs the code
# echo --> display the code
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(plotly)
library(DT)
```

### Introduction 

### Background Information

### Process Overview

What will we be doing? Methods, techniques, why?

### Data Cleaning

A brief look at the raw data can be found below.

```{r, echo = FALSE}
# Import the raw data
majors_raw <- read.csv("/Users/rachael/Documents/School/Fall 2021/DS 3001/Final Project/recent-grads.csv")
DT::datatable(majors_raw)
```
```{r}
str(majors_raw)
```
As can be seen above, many of the categories are integer values. Many of these variables can be converted into factor variables in addition to the numerical ones. In addition, the categorical variable categories can be compressed in order for more useful data for the analysis.

```{r}
# Changing Categorical Variables into Factor Variables


```

### Exploratory Data Analysis

Initial summary statistics and graphs with an emphasis on variables you believe to be important for your analysis. 

```{r}

```

### Data Vizualization

```{r}

```

### Model Building

```{r}

```

### Prediction

```{r}

```

### Evaluation

```{r}

```

### Fairness Assesment

### Conclusion

What can you say about the results of the methods section as it relates to your question given the limitations to your model?

### Future Recommendations

What additional analysis is needed or what limited your analysis on this project?



