---
title: "Final Project"
author: "Rachael Cooper, Sarah Gould, Nathan Patton, Samarth Saxena"
date: "11/29/2021"
output: 
  html_document:
    toc: TRUE
    toc_depth: 5
    toc_float: TRUE
    tod_collapsed: TRUE
    theme: cosmo
  editor_options:
    chunk_output_type: console
    
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# cache --> only have to recompile the changes
# eval --> runs the code
# echo --> display the code
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(plotly)
library(DT)
```

### Introduction 

### Background Information

### Process Overview

What will we be doing? Methods, techniques, why?

### Data Cleaning

```{r}

```

### Exploratory Data Analysis

Initial summary statistics and graphs with an emphasis on variables you believe to be important for your analysis. 

```{r}

```

### Data Vizualization

```{r}

```

### Model Building

```{r}

```

### Prediction

```{r}

```

### Evaluation

```{r}

```

### Fairness Assesment

### Conclusion

What can you say about the results of the methods section as it relates to your question given the limitations to your model?

### Future Recommendations

What additional analysis is needed or what limited your analysis on this project?



